### 2012-05-17 GUIErrorDetector/paper/evaluation.tex
make the call graph construction section shorter, and the paper now fits in 11 pages

We used 10 open-source projects from SourceForge, Google Code,
and the Eclipse plugin marketplace as evaluation subjects.

Five subjects (EclipseRunner, HudsonEclipse, SGTPuzzler, Fennec, and MyTracks)
are selected because they have known invalid thread access errors (1 error per subject) and
all errors have been fixed in later revisions.
We used the buggy versions to check whether our tool
can correctly identify those known errors. 
For the other five subjects, we selected them
by first searching for the
framework keywords (e.g., ``Java Swing'' or ``Java SWT'')
in the above source repositories, and then choosing subjects based on the following
criteria. First, the subject must be a Java application, not
an open library. Second, the subject must use multithreading in its implementation.
Third, the subject is listed in the first 5 result pages.
This permits us to exclude immature subjects that may contain obvious errors.
For each selected subject, we ran our tool on the latest stable release to find new errors.
%of each project to check whether
%it can find any new errors.

@@

We manually determined the validity of each warning.
For a known error, we compared the generated report (i.e., the
error-revealing method call chain as Figure~\ref{fig:report} shows) against the
actual error-fixing patch to check whether our tool correctly identified the
erroneous method. For a previously-unknown error, we submitted a new bug
report to its developers, and wrote a test driver to reproduce it.
%wrote a test driver to

\begin{table*}[ht]
\begin{center}
 \fontsize{9pt}{\baselineskip}\selectfont
\hspace*{-0.2cm}
\setlength{\tabcolsep}{.14\tabcolsep}
\begin{tabular}{|l||c|c|c||c|c|c||c|c|c||c|}
\hline
 Subject&  \multicolumn{9}{|c||}{Our Technique} & Requiring Wrappers  \\
\cline{2-10}
 Program  &  \multicolumn{3}{|c||}{RTA }& \multicolumn{3}{|c||}{0-CFA } & \multicolumn{3}{|c||}{1-CFA } & (Section~\ref{sec:straightforward})  \\
\cline{2-11}
 & CG Size & \#Warning & \#Bug & CG Size & \#Warning & \#Bug & CG Size & \#Warning & \#Bug & \#Warning \\
\hline \hline
\multicolumn{11}{|l|}{SWT desktop applications}   \\
 \hline
 VirgoFTP&  12401 &  1 &  1 & 10858 & 1 & 1 & 43598 & 2 & 2& 149  \\
 \hline
 FileBunker &  18951 &  1 &  0 & 15743 & 0 & 0 & 76088 & 2 & 1& 693  \\
 \hline
 ArecaBackup&  20882 &  1  &  0 & 19697 & 1 & 1  & 116398 & 1 & 1 & 3021\\
 \hline
 \hline
\multicolumn{11}{|l|}{Eclipse plugins}   \\
 \hline
 EclipseRunner&  11248&  6 &  1 & 7201 & 6 & 1 & 26911 & 6 & 1& 202  \\
 %\hline
 %FileSync&  12132 &  xx &  xxx &8235 & xx & xx& 32565 & 1 & 1 & 331 & 1 \\
 \hline
 HudsonEclipse& 18473 &  2 &  1 & 15814 & 2 & 1& 56645 & 3 & 1 & 182 \\
 \hline
 \hline
\multicolumn{11}{|l|}{Swing desktop applications}   \\
 \hline
 S3dropbox & 37751 &  0 &  0 & 30609 & 0 & 0 & 115324 & 1 & 1 & 210  \\
 \hline
  SudokuSolver&  27730&  3 &  2 & 20907 & 3 & 2 & 39299 & 2 & 2 & 356  \\
 \hline
 \hline
\multicolumn{11}{|l|}{Android mobile applications}   \\
 \hline
 SGTPuzzler & 13631 / 13865&  1 / 16 &  0 / 0 & 9546 / 9682& 0 / 4& 0 / 1 & 35198 / 35756 & 0 / 1  & 0 / 1& 104 \\
 \hline
 Fennec & 14058 / 14387 &  1 / 1 &  0 / 0 & 8263 / 8898 & 1 / 1 & 0 / 0& 29125/ 31759 & 3 / 3 & 1 / 1& 433 \\
 \hline
 MyTracks & 24036 / 24036 &  161 / 220 & 0 / 0 & 10803 / 13645 & 119 / 119 & 0 / 0 & 39235 / 110977 & 1 / 1 & 0 / 1 & 1192 \\
\hline
\hline
 Total & 199161 / 199724 &  177 / 251 & 5 / 5 & 149441 / 153054& 133 / 137& 6 / 7 & 577821 / 652756& 21 / 22 & 10 / 12 & 6542\\
\hline
\end{tabular}
\end{center}
\vspace{-15pt}
\Caption{{\label{table:results}Experimental results in finding invalid-thread-access errors
in multithreaded GUI programs. Column ``CG Size'' is the number of nodes in the
call graph. Column ``\#Warning''
is the number of warnings issued by our tool. Column ``\#Bug'' shows
the actual bugs found. Column groups ``RTA'', ``0-CFA'',
and ``1-CFA'' show the results of using different
call graph construction algorithms.
For the Android applications, a slash ``/'' separates the result of
using standard call graph construction algorithm and our call
graph construction algorithm in Section~\ref{sec:cgcon} (dealing
with reflection calls and adding \annotationnum native method annotations
for SGTPuzzler). The \newbugs errors found in VirgoFTP, FileBunker, ArecaBackup,
S3dropbox, and SudokuSolver are previously  unknown. As a comparison, the
results of the ``Requiring Wrappers'' approach (Section~\ref{sec:straightforward}),
are shown at the far right.} }
\end{table*}

@@

As shown in Table~\ref{table:results}, our tool found
errors in each subject. Using the 1-CFA
call graph algorithm, our tool issued 22 warnings, among
which \bugs warnings reveal \bugs distinct errors (\newbugs were previously unknown,
and the \oldbugs known errors were correctly identified),
2 warnings are false positives, and the remaining 8 warnings
are redundant. Section~\ref{sec:reflectionaware} compares with other call graph
construction algorithms.

We submitted all \newbugs new errors
to the respective developers. As of April 2012, 
we have reproduced 3 errors, the S3dropbox developers have confirmed
1 error, and 3 errors in VirgoFTP, FileBunker, and ArecaBackup
have not yet been verified by the developers. All found bugs
and the tool's output are publicly available at: \url{http://www.cs.washington.edu/homes/szhang/guierror/}.

@@

\noindent This variant returns the shortest path from any \CodeIn{Thread.start()} node
to any UI-accessing node, rather than the shortest path from each
\CodeIn{Thread.start()} node to any UI-accessing node as the algorithm in
Figure~\ref{fig:detectalgorithm} does.
This variant reported 10 errors and 12 false positives.   It missed 1 error in S3dropbox and 1 error in
SudokuSolver.
It visits every node in the call graph only once, rather than potentially
once per entry node, but doing so prunes out real error paths.

The second variant uses one DFS per entry node, by changing
the $\mathit{worklist}$ on line 7 of Figure~\ref{fig:detectalgorithm} to
a stack, and changing queue operations dequeue and enqueueAll
on lines 10 and 21
to pop and pushAll.
This variant reported 11 errors and 10 false positives.  It missed 1 error in FileBunker.
DFS tends to search deeper into the graph and return longer paths that are
more likely to be infeasible or to be removed by the filters.
%For the FileBunker subject, paths
%returned by DFS are much longer than BFS, and do not actually exist.

The third variant uses exhaustive search to find potential errors. This variant
enumerates all non-cyclic paths from all reachable \CodeIn{Thread.start()} nodes
to each UI-accessing node, and then checks whether each
path spawns a new thread and accesses GUI objects without using safe UI methods.
However, the number of non-cyclic paths in a graph is exponential in the graph size,
and it is infeasible for this variant to enumerate all paths for a realistic
call graph, which often contains thousands of
nodes (e.g., $2^{1000}$ $\approx$ 1.07$\times$$10^{301}$). In our experiment,
we ran this variant on each subject for 1 hour, it explored 5.1$\times$$10^9$ paths on average, but did not
output any real errors before it terminated.

@@

\noindent \textbf{\textit{Threats to validity.}}
There are two major threats to validity in our evaluation. 
One threat is the degree to which the subject programs
used in our experiment are representative of true practice.
In our evaluation, we only selected subjects from
open-source repositories. Another threat is that we only employed three
widely-used call graph construction algorithms (i.e., RTA, 0-CFA, and 1-CFA) 
in our evaluation. Using other 
call graph construction algorithms such as XTA~\cite{xta} or VTA~\cite{Sundaresan:2000} 
 might achieve different results.

>>>>

We used \subnum open-source projects from SourceForge, Google Code,
and the Eclipse plugin marketplace as evaluation subjects.

Five subjects (EclipseRunner, HudsonEclipse, SGTPuzzler, Fennec, and MyTracks)
are selected because they have known invalid thread access errors (1 error per subject) and
all errors have been fixed in later revisions.
We used the buggy versions to check whether our tool
can correctly identify those known errors. 
For the other four subjects, we selected them
by first searching for the
framework keywords (e.g., ``Java Swing'' or ``Java SWT'')
in the above source repositories, and then choosing subjects based on the following
criteria. First, the subject must be a Java application, not
an open library. Second, the subject must use multithreading in its implementation.
Third, the subject is listed in the first 5 result pages.
This permits us to exclude immature subjects that may contain obvious errors.
For each selected subject, we ran our tool on the latest stable release to find new errors.
%of each project to check whether
%it can find any new errors.

@@

We manually determined the validity of each warning.
For a known error, we compared the generated report (i.e., an
error-revealing method call chain as shown in Figure~\ref{fig:report}) against the
actual bug fix to check whether the tool identified the
buggy method. For a previously-unknown error, we submitted a new bug
report to its developers, and wrote a test driver to reproduce it.
%wrote a test driver to

\begin{table*}[ht]
\begin{center}
 \fontsize{9pt}{\baselineskip}\selectfont
\hspace*{-0.2cm}
\setlength{\tabcolsep}{.14\tabcolsep}
\begin{tabular}{|l||c|c|c||c|c|c||c|c|c||c|}
\hline
 Subject&  \multicolumn{9}{|c||}{Our Technique} & Requiring Wrappers  \\
\cline{2-10}
 Program  &  \multicolumn{3}{|c||}{RTA }& \multicolumn{3}{|c||}{0-CFA } & \multicolumn{3}{|c||}{1-CFA } & (Section~\ref{sec:straightforward})  \\
\cline{2-11}
 & CG Size & \#Warning & \#Bug & CG Size & \#Warning & \#Bug & CG Size & \#Warning & \#Bug & \#Warning \\
\hline \hline
\multicolumn{11}{|l|}{SWT desktop applications}   \\
 %\hline
 %VirgoFTP&  12401 &  1 &  1 & 10858 & 1 & 1 & 43598 & 2 & 2& 149  \\
 \hline
 FileBunker &  18951 &  1 &  0 & 15743 & 0 & 0 & 76088 & 2 & 1& 693  \\
 \hline
 ArecaBackup&  20882 &  1  &  0 & 19697 & 1 & 1  & 116398 & 1 & 1 & 3021\\
 \hline
 \hline
\multicolumn{11}{|l|}{Eclipse plugins}   \\
 \hline
 EclipseRunner&  11248&  6 &  1 & 7201 & 6 & 1 & 26911 & 6 & 1& 202  \\
 %\hline
 %FileSync&  12132 &  xx &  xxx &8235 & xx & xx& 32565 & 1 & 1 & 331 & 1 \\
 \hline
 HudsonEclipse& 18473 &  2 &  1 & 15814 & 2 & 1& 56645 & 3 & 1 & 182 \\
 \hline
 \hline
\multicolumn{11}{|l|}{Swing desktop applications}   \\
 \hline
 S3dropbox & 37751 &  0 &  0 & 30609 & 0 & 0 & 115324 & 1 & 1 & 210  \\
 \hline
  SudokuSolver&  27730&  3 &  2 & 20907 & 3 & 2 & 39299 & 2 & 2 & 356  \\
 \hline
 \hline
\multicolumn{11}{|l|}{Android mobile applications}   \\
 \hline
 SGTPuzzler & 13631 / 13865&  1 / 16 &  0 / 0 & 9546 / 9682& 0 / 4& 0 / 1 & 35198 / 35756 & 0 / 1  & 0 / 1& 104 \\
 \hline
 Fennec & 14058 / 14387 &  1 / 1 &  0 / 0 & 8263 / 8898 & 1 / 1 & 0 / 0& 29125/ 31759 & 3 / 3 & 1 / 1& 433 \\
 \hline
 MyTracks & 24036 / 24036 &  161 / 220 & 0 / 0 & 10803 / 13645 & 119 / 119 & 0 / 0 & 39235 / 110977 & 1 / 1 & 0 / 1 & 1192 \\
\hline
\hline
 Total & 186760 / 187323&  176 / 250 & 4 / 4 & 138583 / 142196& 132 / 136& 5 / 6 & 534223 / 609158& 19 / 20 & 8 / 10 & 6393\\
\hline
\end{tabular}
\end{center}
\vspace{-15pt}
\Caption{{\label{table:results}Experimental results in finding invalid-thread-access errors
in multithreaded GUI programs. Column ``CG Size'' is the number of nodes in the
call graph. Column ``\#Warning''
is the number of warnings issued by our tool. Column ``\#Bug'' shows
the actual bugs found. Column groups ``RTA'', ``0-CFA'',
and ``1-CFA'' show the results of using different
call graph construction algorithms.
For the Android applications, a slash ``/'' separates the result of
using standard call graph construction algorithm and our call
graph construction algorithm in Section~\ref{sec:cgcon} (dealing
with reflection calls and adding \annotationnum native method annotations
for SGTPuzzler). The \newbugs errors found in FileBunker, ArecaBackup,
S3dropbox, and SudokuSolver are previously  unknown. As a comparison, the
results of the ``Requiring Wrappers'' approach (Section~\ref{sec:straightforward}),
are shown at the far right.} }
\end{table*}

@@

As shown in Table~\ref{table:results}, our tool found
errors in each subject. Using the 1-CFA
call graph algorithm, our tool issued \warnings warnings, among
which \bugs warnings reveal \bugs distinct errors (\newbugs were previously unknown,
and the \oldbugs known errors were correctly identified),
2 warnings are false positives, and the remaining 8 warnings
are redundant. Section~\ref{sec:reflectionaware} compares with other call graph
construction algorithms.

We submitted all \newbugs new errors to the respective developers. As of May 2012, 
1 error S3dropbox has been confirmed, and we have reproduced other errors,
in which 2 errors in FileBunker and ArecaBackup
have not yet been verified by the developers. 
%2 errors in SudokuSolver can be reproduced by simply
%clicking its UIs, and the rest 4 errors can be reproduced
%by writing a test driver have not yet been verified
%by the developers.
All found bugs and our experimental results are
publicly available at: \url{http://www.cs.washington.edu/homes/szhang/guierror/}.

@@

\noindent This variant returns the shortest path from any \CodeIn{Thread.start()} node
to any UI-accessing node, rather than the shortest path from each
\CodeIn{Thread.start()} node to any UI-accessing node as the algorithm in
Figure~\ref{fig:detectalgorithm} does.
This variant reported 8 errors and 12 false positives.   It missed 1 error in S3dropbox and 1 error in
SudokuSolver.
It visits every node in the call graph only once, rather than potentially
once per entry node, but doing so prunes out real error paths.

The second variant uses one DFS per entry node, by changing
the $\mathit{worklist}$ on line 7 of Figure~\ref{fig:detectalgorithm} to
a stack, and changing queue operations dequeue and enqueueAll
on lines 10 and 21
to pop and pushAll.
This variant reported 9 errors and 10 false positives.  It missed 1 error in FileBunker.
DFS tends to search deeper into the graph and return longer paths that are
more likely to be infeasible or to be removed by the filters.
%For the FileBunker subject, paths
%returned by DFS are much longer than BFS, and do not actually exist.

The third variant uses exhaustive search to find potential errors. This variant
enumerates all non-cyclic paths from all reachable \CodeIn{Thread.start()} nodes
to each UI-accessing node, and then checks whether each
path spawns a new thread and accesses GUI objects without using safe UI methods.
We ran this variant on each subject for 1 hour, it explored 5.1$\times$$10^9$ paths
on average, but did not output any real errors before it terminated.
The number of non-cyclic paths in a graph is exponential in the graph size,
and it is infeasible to enumerate all paths for a realistic call graph.
In our experiments, the smallest call graph contains 7201 nodes and the average out-degree
of each node is 2.15. Thus, a rough estimation for the number of distinct
non-cyclic paths leads to an astronomically large number: 
$2.15^{7201}$ $\approx$ 1.07$\times$$10^{2107}$.

@@

\noindent \textbf{\textit{Threats to validity.}}
There are two major threats to validity in our evaluation. 
One threat is the degree to which the subject programs
used in our experiment are representative of true practice.
In our evaluation, we only selected subjects from
open-source repositories. Another threat is that we only employed three
widely-used call graph construction algorithms (i.e., RTA, 0-CFA, and 1-CFA) 
in our evaluation. Using other 
call graph construction algorithms %such as XTA~\cite{xta} or VTA~\cite{Sundaresan:2000} 
 might achieve different results.
